### 2022-01-30 main.tex
add anynomous

\documentclass{article}
\usepackage[english]{babel}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Anonymous provision of services via blockchain}
% Anonymous and fair provision of physical services via blockchain
\author{Stanisław Barański}
\date{December 2021}

\providecommand{\keywords}[1]{\textbf{Keywords:} #1}

\begin{document}
\maketitle

\begin{abstract}
Service providers like: lawyers, laboratories, auditors, or banks, to provide services, require customers to submit data that is often associated with their personal information. This situation expose customers to privacy risk. However, most of the service providers use the personal information merely for logistic operations, and could provide services on anonymous data if other means of communication would be possible.

We propose a protocol for local confidential services provision, using blockchain for both contract agreement and anonymous payments.

In case of dispute, we let the parties communicate over blockchain. In case of disagreement, the customer can disclose the whole interaction and proof to the justice misbehaving SP.

We propose three applications for our protocol: (i) anonymous steroid examination, (ii) anonymous legal advice, and (iii) anonymous financial trustworthiness evaluation.

\end{abstract}
\keywords{data privacy, privacy preserving, blockchain, fair-exchange}

\section{Problem statement}
Service providers (SP) like: lawyers, laboratories, auditors, or banks, to provide services, require data that is often associated with user identity.

Providing personal information exposes users to privacy risk, i.e. potential loss of control over personal information ~\cite{smith2011information}.

Such user information can be then used deliberately or unintentionally (e.g, by theft) for: insider disclosure, unauthorized access, or commercial gains. For example, by reselling it to marketers, financial institutions, other businesses, government agencies, or even cybercriminals. Which in turn can lead to profiled advertisements, or criminal activities like identity theft or illegal tracking and surveillance~\cite{smith2011information}.

The guarantee of the privacy of the data is based on a trust assumptions and security of IT systems. However, most of the service providers don't need the information of user identity for other reasons than payment, communication, or logistics.

It would be desirable if the user could keep its identity private, while service provider still provide its services. This could lead to reduced trust that users have to put on SP and less responsibility borne by SP.

In this paper we propose a protocol for anonymous provision of services via blockchain and cryptography techniques. Specifically, we use
\begin{enumerate}
    \item \textbf{confidential blockchain} to process payment for a service without revealing a user's identity.
    \item \textbf{smart contracts} to process the agreements and disputes in decentralized and trustless manner.
    \item \textbf{zero-knowledge proofs} — to proof the payment for the service without revealing the payment itself.
    \item \textbf{encryption scheme and blockchain} — to asynchronously deliver the results of the made service back to the user.
\end{enumerate}

Authors of the paper~\cite{lin2021efficient} propose a system for Privacy-Preserving Credit Score System. Letting banks to calculate creditworthness without learning user private information like registration, hobbies, credit, relationships, and inquiry. To achieve this they use Paillier encryption and noninteractive zero-Knowledge schemes (NIZK).

\subsection{Fair-exchange protocols}

We observe that such protocols for providing services are special case of fair-exchange protocol.

Fair-exchange protocol is a protocol used for performing e-commerce transactions of both digital and physical goods, exchanging documents, or signing contracts, between mutually distrusting parties, while satisfying following conditions:
\begin{itemize}
    \item non-repudiation  — any participating party can not deny their actions (e.g. signing, sending, and receiving data). Usually this is achieved by digital signatures.
    \item fairness — any participating party does not gain any advantage at any step of the exchange protocol.
    \item timeliness — the protocol eventually terminates for all participants.
\end{itemize}

There are two main categories of fair-exchange protocols~\cite{kremer2002intensive,duangphasuk2020fair}: those that involve the use of trusted third party (TTP), and those that do not.

\paragraph{Protocols that do not involve use of trusted third party.}
Protocols that do not involve use of trusted third party, also called "probabilistic" or "gradual" exchange protocols. In such protocols~\cite{markowitch1999probabilistic}, items that will be exchanged are first split into parts. Then, the parts are exchanged between parties. This causes increased communication overhead. Moreover the fairness of such protocol is probabilistic, meaning that there is a small chance that the protocol will end up with one party being in advantageous position.

Hence, the great benefit of TTP-less protocol, comes at a cost of increased communication overhead, and unsatisfactory property of uncertain termination.

\paragraph{Protocols that involve the use of trusted third party}
The most common approach to achieve fair-exhange conditions is by introducing Trusted Third Party (TTP).

TTP can play different role in fair-exchange protocol. Commonly agreed classification~\cite{kremer2002intensive,duangphasuk2020fair} distinguish three roles the TTP can play:

\begin{enumerate}
    \item Inline~\cite{coffey1996non}, the TTP act as a authoritative middleman between sender and receiver. The obvious consequence of such method is the communication bottleneck and single point of failure. These protocols are inefficient and trust-dependent.

    \item Online~\cite{djuric2015feips}, the TTP is involved in each protocol step, but not in each data transmission. Significantly improving the performance compared to inline TTP.

    \item Offline~\cite{hwang2015provable}, the TTP is involved only in case of a dispute between parties. This approach is also called "optimistic", because in scenarios where both parties behaves honesty, the TTP is ignored. Resulting in high-performance and low trust-dependency. 
\end{enumerate}

Involving online and centralized TTP in each transaction, could cause the communication bottleneck problem. Therefore, the offline TTP is the most favorable approach.

The main idea of the offline TTP model is that the sender sends a ciphertext of his/her own document before obtaining the expect document from opposite party. Both parties will then fairly exchange the decryption keys of the ciphertext.

\section{E-commerce transaction protocol}

According to ~\cite{djuric2015feips}, a typical e-commerce process can be illustrated with the diagram on figure ~\ref{fig:ecommerce-transaction-diagram}.

\begin{figure}[htbp]
\centerline{\includegraphics[width=.4\linewidth]{ecommerce-process.png}}
\caption{Diagram illustrating a typical e-commerce transaction process}
\label{fig:ecommerce-transaction-diagram}
\end{figure}

\section{Pseudo-anonymous vs anonymous blockchain}
In the papers on fair-exchange and anonymity, there's a disagreement on what the anonymity is, the definition is vague, and sometimes overused (wiki leaks), it's rather commonly accepted that Bitcoin and other the like traceable blockchains are not anonymous, rather pseudo-nomous.

What is a anonymity then? We like the definition stated in (huge paper on privacy ), or the more practical in the context of blockchain the definition stated in (swap in Monero).

\section{State of the art}

\subsection{Anonymous fair-exchange}

An protocol ensures anonymity if no party can make a link between identity and actions taken by another party.

\paragraph{Themis} Themis~\cite{meng2019themis} is a fair-exchange protocol, which uses blockchain instead of TTP. It provide escrow services for secure exchanging between cryptocurrencies and digital goods.

Authors highlight that a widely-used atomic exchange mechanism HTLC (Hashed Time-lock Contract) is vulnerable to denial of service (DoS) attacks. While Multi-Signature Schemes depends on the Trusted Third Party (TTP) making them vulnerable to collusion attacks. Such TTP can collude with one party and take the locked funds. It can also refuse settling disputes and lock the hosted funds.

Themis is a decentralized system that provides an escrow and dispute resolution. It claims to satisfy following requirements:
\begin{description}
    \item[Fairness] Either both seller and buyer obtain all the goods, or they both obtain nothing.
    \item[Security] The funds are locked until the transaction is completed.
    \item[Passivity] If no dispute arise, the third-party does not need to involve in the transaction.
    \item[Correctness] Transactions and disputes are executed according to a priori agreed protocol rules.
    \item[Dependability] There is no single point of failure vulnerable to DoS attack.
    \item[Privacy] If the dispute does not arise, the third-party does not know if the transaction completes. If the dispute arises, only the related parties are aware of it.    
\end{description}

Two parties, say Alice and Bob, can create an escrow by generating 2-of-2 (not 2-of-3 as it is other TTP-based systems) threshold account using Thresh-Key-Gen protocol. Alice and Bob generate secret keys $x_A$ and $x_B$ accordingly, the shared account public key becomes $y = g^{x_A+x_B}$; only the access to both Alice's and Bob's secrets grants access to the shared account.

Then, both Alice and Bob, takes their keys, and split them into $n=2t+1$ secret shares using Shamir Secret Shamir protocol, where $n$ is the number of mediators that participate in the decentralized network, and $t+1$ becomes the threshold of sufficient number of mediators that can reconstruct the secret key.

Next, both Alice and Bob, take their key shares, and encrypt each $i$-th key share using public key of $i$-th mediator. As a result Alice's secret $x_A$ becomes ${c^A_1, c^A_2,...,c^A_n}$, and Bob's secret $x_B$ becomes ${c^B_1, c^B_2,...,c^B_n}$, where $c_i = E_{M_i}(P_i)$ is the encrypted $i$-th key share using $i$-th mediator public key $M_i$.

Next, both Alice and Bob exchange the sets of encrypted key shares with each other. And send funds to the escrow account.

The escrow is secure as long as $t+1$ of mediators does not collude, which would let them recreate both $x_A$ and $x_B$. Also, to ensure that parties are exchanging real key shares, they also send a witnesses generated with Feldman VSS scheme and zero-knowledge proofs to guarantee the consistency between witnesses and key shares.

In case of dispute, the decentralized network of mediators settle the conflict and grand one winning party, the other party secret key, letting him to withdraw the funds.

The honesty of mediators is achieved by the monetary incentivization and reputation system.

\paragraph{A Solution for Secure Certified Electronic Mail Using Blockchain as a Secure Message Board}

The idea of certified mail is simple, neither the sender can deny sending the mail, nor the receiver can deny receiving it. As the service is widely used in paper world, achieving it for e-mails haven't meet the agreement across scientific community for over 40 years. The main reason was the dependency of TTP. Which significantly reduce the performance, security, and robustness of such protocol. Those protocols that do not use TTP struggle with high computation and communication overhead.

Authors of ~\cite{hinarejos2019solution} replaced the TTP with blockchain (concretely Bitcoin blockchain as a reference implementation), which can be seen as a secure, verifiable, and decentralized TTP.

The idea behind certified e-mails—and any other fair-exchange protocol—is following
\begin{enumerate}
    \item The sender sends encrypted and signed message to the receiver.
    \item The receiver returns a proof of delivery (a signature) of the encrypted message to the sender.
    \item The sender publish on blockchain (or TTP in general) a proof of delivery and the decryption key.
    \item The receiver decrypt the encrypted message using the published decryption key.
\end{enumerate}

The non-repudiation requirement is achieved by the receiver sending the Proof-of-Delivery prior having access to the decrypted message. And the sender publishing the decription key together with the Proof-of-Delivery on blockchain (or TTP) so that the receiver can not deny both receiving the message and having access to decryption key, because it's publicly available.

The protocol also handles confidentiality — the encryption key used to encrypt and then decrypt the message is encrypted using the receiver's public key, so the key is never published in raw form.
Also, the protocol supports timeliness, meaning that the transaction is always finalized in a finite time.

In this case, the role of Blockchain is to attest the existence of decryption key.

\subsection{Anonymous delivery system for physical goods}

\paragraph{Anonymity and fair-exchange in e-commerce protocol for physical products delivery}

Authors of ~\cite{birjoveanu2015anonymity} proposed a fair-exchange of physical product and electronic payment protocol that guarantee both customer and merchant anonymity. They achieve it by introducing online trusted third party that validate coins and provide fair exchange guarantees.

The anonymity is ensured by assuming the existence of source cabinet (SC) and delivery cabinet (DC), and a trusted delivery agent who take product from a source cabinet and provide it to a destination cabinet. 
Both cabinets provides access to the product by passwords to conceal identity of the customer and the merchant.
To avoid identity disclosure of customer and merchant, the access to the product is provided by passwords.

Both cabinets are equipment with cameras. Also there should exist the trusted delivery agent (DA) who will deliver items from source cabinet do delivery cabinet.

The protocol works as follows:
\begingroup
\renewcommand{\labelenumii}{\arabic{enumii}.}
\begin{enumerate}
    \item Both parties agree on the purchase details.
    \item The customer buys a digital coin from his bank and validates it with the TTP.
    \item The customer sends to the merchant the purchase order and the digital signature made by TTP on the encrypted coin.
    \item The merchant post the product to the source cabinet.
    \item Delivery agent collect the product from source cabinet and post it to delivery cabinet.
    \item Customer collects the product from delivery cabinet using password.
    \item The customer checks if the product is actually the ordered one.
    \begin{itemize}
    \item[-] if yes 
        \begin{enumerate}
        \setcounter{enumii}{7}
        \item the customer sends the acknowledge to delivery cabinet and decryption key to the merchant.
        \item The merchant redeem the coins from the bank.
        \end{enumerate}
    \item[-] if no
        \begin{enumerate}
        \setcounter{enumii}{7}
        \item the delivery cabinet is equipment with video camera that records the moment when the customer open the package and let the customer to signal the invalidity of the product, and so, sending the encrypted recording to the TTP. 
        \item the dispute is settled via TTP, optionally letting each party to reveal it's identity.
    \end{enumerate}
    \end{itemize}
\end{enumerate}
\endgroup

Yet, the fair-exchange properties are achieved by strong assumptions. Namely, 
\begin{itemize}
    \item the TTP does not misbehave or collude with any party. 
    \item The customer's and merchant's banks enable confidental transactions and both share commit-buffer where the value is locked until the transaction is finished.   
    \item All banks mantain a global list of coin's serial numbers to prevent double-spending problem. 
    \item The source cabinet and delivery cabinet exists, and in both the product access is protected by password.
    \item Delivery cabinet is equppted with a video camera that records the moment when the customer opens the package, and provide a way to submit the video to TTP in case of a dispute. 
    \item Delivery agent exist and is trusted. 
    \item Communication channels between parties provides anonymity.
\end{itemize}

\paragraph{Lelanros} Lelanros~\cite{altawy2017lelantos} is a blockchain-based anonymous physical delivery system. The anonymity is achieved by employing ToR's onion routing to connect physical delivery providers. The whole route from merchant to customer is split into multiple steps, and each step is undertaken by different randomly selected delivery provider. As long as the develiery providers does no collude, mechant can not learn both identity and destination address of the customer.

Onchain smart-contract is used to coordinate the whole process and mediate communication between customer and delivery providers.

However, the system achieve only pseudo-anonymity or 'weak anonymity'.

\paragraph{Auditable Protocols for Fair Payment and Physical Asset Delivery Based on Smart Contracts}
Authors of~\cite{wang2019auditable} propose simillar to \cite{altawy2017lelantos} physical asset delivery system to connect merchants, customers and logistic company.

However, the system does not achieve anonymity.

\section{Contribution}
In this paper similarly to by~\cite{birjoveanu2015anonymity}, we propose fully-fledged protocol for anonymous fair-exchange which includes payment processing and physical goods exchange. Yet, compared to ~\cite{birjoveanu2015anonymity}, we use blockchain instead of centralized TTP. Additionally, we use anonymous cryptocurrencies getting rid of strong and unrealistic assumptions on traditional banking.

Although there have been number of protocols achieving fair-exchange over blockchain~\cite{heilman2016blindly, meng2019themis, icete18, barber2012bitter}, only few achieve anonymity~\cite{heilman2016blindly}. From these achieving anonymity, non target exchange of psychical goods.

We contribute to the field of anonymous fair-exchange and anonymous physical delivery with following improvements:
With this paper we propose a protocol that add following contribution to the field of physical delivery, fair-exchange protocols:
\begin{itemize}
    \item Strong anonymity — unlike other protocols based on pseudoanonymous traceable blockchains like Bitcoin or Ethereum, we propose a solution based on private blockchain (ZCash) where transactions can not be analyzed and thus linked back to users' identity.
    \item Practicality — out proposal is practical and ready to deploy
\end{itemize}

In this paper, we propose a protocol for provisioning local services, that guarantee following requirements:
\begin{enumerate}
    \item Anonymous — The customer identity is strongly untraceable (see section~\ref{anonimity-vs-pseudoanonimity}.
    \item Fair-exchange — the service provider does not receive the payment before it provide the service, in case it provide wrong service, customer can proof this case to the justice.
    \item Blockchain-based — the protocol does not use TTP, instead it uses decentralized blockchain network.
    \item Timeliness — the transaction will terminate if it don't meet the timely requirerement
\end{enumerate}

Additionally, since the services are provided locally, we enable customers to disclose the transaction and start the dispute in law.

Once the package is delivered to the SP, the SP publish POR (Proof-of-recipment) on the blockchain. From this point the customer is secured, i.e. if the merchant misbehaves, customer can proof to the justice that he paid for the service at a particular point of time, and that SP received the package, but didn't publish results.

Or we can create a smart contract that automatically punish the SP if he doesn't fulfull the agreement in time. And the proof to justice is used in case where the content of the service is invalid, something that have to be settled by human.

If merchant misbehaves i.e. does not provide the agreed services, then the customer can create a proof that such incident happened 
However it provide a proof that such incident happened. In settings where the identity of the party is known, such proof can be used to bring the misbehaving party to the justice.

\section{Applications}
\subsection{Anonymous clinical tests}
Bodybuilders, athletes,  weightlifters, cyclists, football and rugby players have to pass anti-doping tests before attending the competition.

Patients willing to check steroid tests, drug tests, venereal diseases — anything that they wouldn't want to share with anyone, have to risk that the data stored in laboratory won't get to unauthorized hands like attacker or malicious worker.

Analysis of urine, saliva, stool, or blood.

\subsection{Anonymous lawyer consultations}

People willing to take business action that may be risky or not yet fully legistlated, may be willing to receive the estimation of risks and repercussions of such action.
Such protocol would provide a mechanism for

\subsection{Anonymous financial trustworthiness evaluation}

Evaluating financial trustworthiness is required when user want to, for example issue a credit card, take loan for a home, or rent a car. 
As defined in~\cite{lin2021efficient}, such a system consist of three parties: (i) users, (ii) credit bureaus, and (iii) creditors. Users willing to take a credit, have to provide to the credit bureaus, their financial activities like: new account creation, account balance, credit card utilization, credit inquires, and payment history. Such information are then submitted to creditors to compute the user's credit score.

\subsection{Anonymous physical delivery e-commerce}
% TODO write about
~\cite{birjoveanu2015anonymity}

% [![](https://mermaid.ink/img/eyJjb2RlIjoic2VxdWVuY2VEaWFncmFtXG4gICAgQ3VzdG9tZXItPj5TbWFydCBDb250cmFjdDogUHJvdmlkZSBkZXRhaWxzLCBsb2NrIGZ1bmRzLlxuXHRhY3RpdmF0ZSBTbWFydCBDb250cmFjdFxuXHRTbWFydCBDb250cmFjdCAtPj4gU21hcnQgQ29udHJhY3Q6IENyZWF0ZSBhZ3JlZW1lbnRcblx0ZGVhY3RpdmF0ZSBTbWFydCBDb250cmFjdFxuXHRhbHQgSWYgc2VydmljZSBwcm92aWRlciBkZWNsaW5lIGFncmVlbWVudFxuXHRTZXJ2aWNlIFByb3ZpZGVyIC0-PiBTbWFydCBDb250cmFjdDogRGVjbGluZVxuXHRTbWFydCBDb250cmFjdCAtPj4gQ3VzdG9tZXI6IFJlZnVuZFxuXHRlbmRcblx0U2VydmljZSBQcm92aWRlciAtPj4gU21hcnQgQ29udHJhY3Q6IEFjY2VwdFxuXHRcblx0YWx0IGRpZ2l0YWwgZG9jdW1lbnRzXG5cdFx0cGFyIGF0b21pYyB0cmFuc2FjdGlvblxuXHRcdEN1c3RvbWVyIC0-PiAgb2ZmLWNoYWluIHN0b3JhZ2U6IFB1Ymxpc2ggZW5jcnlwdGVkIGRvY3VtZW50c1xuXHRcdG9mZi1jaGFpbiBzdG9yYWdlIC0tPj4gQ3VzdG9tZXI6IFByb29mLW9mLURlbGl2ZXJ5IChQT0QpXG5cdFx0ZW5kXG5cdEN1c3RvbWVyIC0-PiBTbWFydCBDb250cmFjdDogUHVibGlzaCBQT0Rcblx0ZWxzZSBwaHlzaWNhbCBkb2N1bWVudHNcblx0XHRwYXIgYXRvbWljIHRyYW5zYWN0aW9uXG5cdFx0Q3VzdG9tZXIgLT4-ICBTZXJ2aWNlIFByb3ZpZGVyOiBQcm92aWRlIGRvY3VtZW50cyBwaHlzaWNhbGx5XG5cdFx0U2VydmljZSBQcm92aWRlciAtPj4gU21hcnQgQ29udHJhY3Q6IFB1Ymxpc2ggUE9EXG5cdFx0ZW5kXG5cdGVuZFxuXHRsb29wIGV4ZWN1dGUgdW50aWwgY3VzdG9tZXIgaXMgc2F0aXNmaWVkIHdpdGggcmVzdWx0c1xuXHRTZXJ2aWNlIFByb3ZpZGVyIC0-PiBTZXJ2aWNlIFByb3ZpZGVyOiBFeGVjdXRlIHNlcnZpY2Vcblx0U2VydmljZSBQcm92aWRlciAtPj4gT2ZmLWNoYWluIHN0b3JhZ2U6IFB1Ymxpc2ggZW5jcnlwdGVkIHJlc3VsdHNcblx0U2VydmljZSBQcm92aWRlciAtPj4gU21hcnQgQ29udHJhY3Q6IFB1Ymxpc2ggUHJvb2Ytb2YtUmVzdWx0c1xuXHRcblx0XG5cdEN1c3RvbWVyIC0-PiBTbWFydCBDb250cmFjdDogR2V0IFByb29mLW9mLVJlc3VsdHNcblx0Q3VzdG9tZXIgLT4-IE9mZi1jaGFpbiBzdG9yYWdlOiBHZXQgcmVzdWx0c1xuXHRDdXN0b21lciAtPj4gQ3VzdG9tZXI6IFZhbGlkYXRlIHJlc3VsdHNcblx0YWx0IHJlc3VsdHMgYXJlIGluY29ycmVjdFxuXHRcdEN1c3RvbWVyIC0-PiBTbWFydCBDb250cmFjdDogU3RhcnQgZGlzcHV0ZSwgcHJvdmlkZSBkZXRhaWxzXG5cdFx0YWx0IFNlcnZpY2UgUHJvdmlkZXIgZGVjbGluZSBhcmd1bWVudHNcblx0XHRcdFNtYXJ0IENvbnRyYWN0IC0-PiBTZXJ2aWNlIFByb3ZpZGVyOiBVbmxvY2sgZnVuZHNcblx0XHRcdG5vdGUgcmlnaHQgb2YgU2VydmljZSBQcm92aWRlcjogU1AgaXMgYWx3YXlzIG9uIGxvc3QgcG9zaXRpb24sIDxicj5iZWNhdXNlIGhlIGNhbiBub3Qgc3VlIGFub255bW91cyBjdXN0b21lci48YnI-IFRoZXJlZm9yZSB0aGUgc2hvdWxkIGJlIHRoZSBob2xkZXIgdGhlIGZ1bmRzIGFzIHNvb24gYXMgcG9zc2libGUuXG5cdFx0XHRDdXN0b21lciAtPj4gUG9saWNlOiBPcGVuIGRpc3B1dGUgYXQgcG9saWNlXG5cdFx0ZW5kXG5cdGVuZFxuXHRlbmRcblx0XG5cbiAgICBcbiAgICAgICAgICAgICIsIm1lcm1haWQiOnsidGhlbWUiOiJkZWZhdWx0In0sInVwZGF0ZUVkaXRvciI6ZmFsc2UsImF1dG9TeW5jIjp0cnVlLCJ1cGRhdGVEaWFncmFtIjpmYWxzZX0)](https://mermaid.live/edit#eyJjb2RlIjoic2VxdWVuY2VEaWFncmFtXG4gICAgQ3VzdG9tZXItPj5TbWFydCBDb250cmFjdDogUHJvdmlkZSBkZXRhaWxzLCBsb2NrIGZ1bmRzLlxuXHRhY3RpdmF0ZSBTbWFydCBDb250cmFjdFxuXHRTbWFydCBDb250cmFjdCAtPj4gU21hcnQgQ29udHJhY3Q6IENyZWF0ZSBhZ3JlZW1lbnRcblx0ZGVhY3RpdmF0ZSBTbWFydCBDb250cmFjdFxuXHRhbHQgSWYgc2VydmljZSBwcm92aWRlciBkZWNsaW5lIGFncmVlbWVudFxuXHRTZXJ2aWNlIFByb3ZpZGVyIC0-PiBTbWFydCBDb250cmFjdDogRGVjbGluZVxuXHRTbWFydCBDb250cmFjdCAtPj4gQ3VzdG9tZXI6IFJlZnVuZFxuXHRlbmRcblx0U2VydmljZSBQcm92aWRlciAtPj4gU21hcnQgQ29udHJhY3Q6IEFjY2VwdFxuXHRcblx0YWx0IGRpZ2l0YWwgZG9jdW1lbnRzXG5cdFx0cGFyIGF0b21pYyB0cmFuc2FjdGlvblxuXHRcdEN1c3RvbWVyIC0-PiAgb2ZmLWNoYWluIHN0b3JhZ2U6IFB1Ymxpc2ggZW5jcnlwdGVkIGRvY3VtZW50c1xuXHRcdG9mZi1jaGFpbiBzdG9yYWdlIC0tPj4gQ3VzdG9tZXI6IFByb29mLW9mLURlbGl2ZXJ5IChQT0QpXG5cdFx0ZW5kXG5cdEN1c3RvbWVyIC0-PiBTbWFydCBDb250cmFjdDogUHVibGlzaCBQT0Rcblx0ZWxzZSBwaHlzaWNhbCBkb2N1bWVudHNcblx0XHRwYXIgYXRvbWljIHRyYW5zYWN0aW9uXG5cdFx0Q3VzdG9tZXIgLT4-ICBTZXJ2aWNlIFByb3ZpZGVyOiBQcm92aWRlIGRvY3VtZW50cyBwaHlzaWNhbGx5XG5cdFx0U2VydmljZSBQcm92aWRlciAtPj4gU21hcnQgQ29udHJhY3Q6IFB1Ymxpc2ggUE9EXG5cdFx0ZW5kXG5cdGVuZFxuXHRsb29wIGV4ZWN1dGUgdW50aWwgY3VzdG9tZXIgaXMgc2F0aXNmaWVkIHdpdGggcmVzdWx0c1xuXHRTZXJ2aWNlIFByb3ZpZGVyIC0-PiBTZXJ2aWNlIFByb3ZpZGVyOiBFeGVjdXRlIHNlcnZpY2Vcblx0U2VydmljZSBQcm92aWRlciAtPj4gT2ZmLWNoYWluIHN0b3JhZ2U6IFB1Ymxpc2ggZW5jcnlwdGVkIHJlc3VsdHNcblx0U2VydmljZSBQcm92aWRlciAtPj4gU21hcnQgQ29udHJhY3Q6IFB1Ymxpc2ggUHJvb2Ytb2YtUmVzdWx0c1xuXHRcblx0XG5cdEN1c3RvbWVyIC0-PiBTbWFydCBDb250cmFjdDogR2V0IFByb29mLW9mLVJlc3VsdHNcblx0Q3VzdG9tZXIgLT4-IE9mZi1jaGFpbiBzdG9yYWdlOiBHZXQgcmVzdWx0c1xuXHRDdXN0b21lciAtPj4gQ3VzdG9tZXI6IFZhbGlkYXRlIHJlc3VsdHNcblx0YWx0IHJlc3VsdHMgYXJlIGluY29ycmVjdFxuXHRcdEN1c3RvbWVyIC0-PiBTbWFydCBDb250cmFjdDogU3RhcnQgZGlzcHV0ZSwgcHJvdmlkZSBkZXRhaWxzXG5cdFx0YWx0IFNlcnZpY2UgUHJvdmlkZXIgZGVjbGluZSBhcmd1bWVudHNcblx0XHRcdFNtYXJ0IENvbnRyYWN0IC0-PiBTZXJ2aWNlIFByb3ZpZGVyOiBVbmxvY2sgZnVuZHNcblx0XHRcdG5vdGUgcmlnaHQgb2YgU2VydmljZSBQcm92aWRlcjogU1AgaXMgYWx3YXlzIG9uIGxvc3QgcG9zaXRpb24sIDxicj5iZWNhdXNlIGhlIGNhbiBub3Qgc3VlIGFub255bW91cyBjdXN0b21lci48YnI-IFRoZXJlZm9yZSB0aGUgc2hvdWxkIGJlIHRoZSBob2xkZXIgdGhlIGZ1bmRzIGFzIHNvb24gYXMgcG9zc2libGUuXG5cdFx0XHRDdXN0b21lciAtPj4gUG9saWNlOiBPcGVuIGRpc3B1dGUgYXQgcG9saWNlXG5cdFx0ZW5kXG5cdGVuZFxuXHRlbmRcblx0XG5cbiAgICBcbiAgICAgICAgICAgICIsIm1lcm1haWQiOiJ7XG4gIFwidGhlbWVcIjogXCJkZWZhdWx0XCJcbn0iLCJ1cGRhdGVFZGl0b3IiOmZhbHNlLCJhdXRvU3luYyI6dHJ1ZSwidXBkYXRlRGlhZ3JhbSI6ZmFsc2V9)
\begin{figure}[htbp]
\centerline{\includegraphics[width=\linewidth]{sequence-flow.png}}
\caption{Sequence flow}
\label{fig:sequence-flow}
\end{figure}

\section{Solution protocol}

\section{confidential Blockchains}

\paragraph{Ethereum}

\paragraph{Aleo}

\paragraph{STARK-net}

\paragraph{Monero}
Monero offers out-of-the-box proving/checking confidental transactions. \href{https://www.getmonero.org/resources/user-guides/prove-payment.html}

\paragraph{ZCash}
ZCash offers payment disclosure as a preview feature.
\href{https://garethtdavies.medium.com/an-introduction-to-payment-disclosure-in-zcash-96748c209d49}

\href{https://garethtdavies.medium.com/an-introduction-to-payment-disclosure-in-zcash-96748c209d49}

\section{Future work}

Verifiable Document Redacting ~\cite{chabanne2017verifiable} is a protocol for erasing confidential information from authenticated images in verifiable manner. In other words, the protocol allows verifying that only the fields that cover personal information has been blacked out from the authenticated image. This allows users to provide required images without exposing their personal information, and SPs to authenticate the modified images, such that they are sure that only the stated fields have been blacked out and nothing else was touched.

If such Verifiable Redacted Documents were accepted by other service providers like bank or tax office, national register of the judiciary, then the SP could use the documents provided by the customer to provide wider range of services, without knowing identity of the customer.

% In this paper we extend the idea to a whole user-to-service-provider protocol

\bibliographystyle{alpha}
\bibliography{bibliography}

\end{document}