### 2023-02-01 paper/highlights.tex
prepare els paper docs

\documentclass[preprint,3p]{elsarticle}

\input{constants.tex}
\begin{document}
\begin{titlepage}
   \begin{center}
       \vspace*{1cm}

       \begin{enumerate}
        \item We design a protocol for anonymous provision of privacy-sensitive services
        \item Privacy guarantee was achieved using anonymous payment methods
        \item We implemented demonstration of the practical application of the protocol
        \item The fairness of the protocol was proven
        \item The source code of the protocol has been made publicaly available
       \end{enumerate}

   \end{center}
\end{titlepage}
\end{document}