### 2023-03-21 paper/contributions.tex
add contributions, coverletter and biographies

\documentclass[preprint,3p]{elsarticle}

\input{constants.tex}
\begin{document}
{\Large Contributions \par}

\vspace{2cm}

\textbf{\STANLONG}: Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Resources, Data Curation, Writing - Original Draft, Writing - Review \& Editing, Visualization

\vspace{1cm}

\textbf{\JULLONG}: Supervision, Project administration, Funding acquisition, Writing - Original Draft, Writing - Review \& Editing

\vspace{1cm}

\textbf{\HIGILONG}: Validation, Writing - Review \& Editing, Methodology

\end{document}