### 2023-03-21 paper/coverletter.tex

\documentclass[preprint,3p]{elsarticle}

\input{constants.tex}
\begin{document}

Dear IEEE Access Editorial Board,

I am writing to submit my manuscript titled "\TITLE" for consideration in IEEE Access. The manuscript proposes a solution that enables service providers to offer sensitive services, such as legal advice, genetic testing, or STD tests, without collecting personal data, and therefore without violating data privacy regulations.

The proposed protocol uses blockchain technology to achieve fairness and anonymity, anonymous payment methods such as cash and cryptocurrencies, and distributed content-addressable storage networks to deliver results. The manuscript compares the proposed protocol with existing solutions and demonstrates its advantages, including weaker assumptions for anonymity, support for physical materials, and practicality.

The main contributions of the paper are as follows:
\begin{enumerate}
    \item We design a protocol for anonymous provision of privacy-sensitive services
    \item Privacy guarantee was achieved using anonymous payment methods
    \item We implemented demonstration of the practical application of the protocol
    \item The fairness of the protocol was proven
    \item The source code of the protocol has been made publicaly available
\end{enumerate}

Thank you for your time and consideration. I look forward to the opportunity to share my work with the IEEE Access community and contribute to the advancement of the field.

\noindent Sincerely,

\noindent Stanislaw Baranski

\end{document}