### 2023-03-21 paper/higinio-bio.tex

\documentclass[preprint,3p]{elsarticle}

\input{constants.tex}
\begin{document}
{Higinio Mora} received the B.S. degree in computer science engineering and the B.S. degree in business studies from the University of Alicante, Spain, in 1996 and 1997, respectively, and the Ph.D. degree in computer science from the University of Alicante in 2003. Since 2002, he has a member of the faculty of the Computer Technology and Computation Department, University of Alicante, where he is currently an Associate Professor and a Researcher with the Specialized Processors Architecture Laboratory. His research interests include computer modeling, computer architectures, high-performance computing, embedded systems, the Internet of Things, and cloud computing paradigm.
\end{document}