### 2023-03-21 paper/julian-bio.tex

\documentclass[preprint,3p]{elsarticle}

\input{constants.tex}
\begin{document}
{Julian Szyma{\'n}ski}  received the PhD degree in 2009 and DSc in 2020 at Gdańsk University of Technology.
He is currently with the Department of Computer
Architecture, Faculty of Electronics, Telecommunications and Informatics, Gdansk University of Technology. He addresses the problems of knowledge representation, methods of lexical knowledge acquisition, and linguistic data utilization.
His research results find applications in web search engines and systems of automated text categorization. He manages research projects for processing the information in Wikipedia. His mainstream of this research is building semantic memory models that improve natural language processing.  
\end{document}